---
title: "censusClean"
output: html_notebook
---

This file cleans raw census St. Louis tract level census data from the 2016 ACS five-year estimates and the 1970 Brown LTDB data. It will combine both data sources to create a single dataset that will be merged into a 2016 tract level shapefile for St. Louis City. 

```{r}
library(brownLTDB)
library(tidyr)
library(dplyr)
library(sf)
```

Import raw 2016 ACS data and brown data from the LTDB libary. 

```{r}
acs2016 <- read.csv("/Users/thbraswell/Documents/Local/barriersCrime/rawData/barriersRawACS.csv", header = TRUE)
full1970 <- ltdb_stl(year = 1970, type = "full")
sample1970 <- ltdb_stl(year = 1970, type = "sample")
```

Wrangle ACS data. Note: raw data dictionary located in "barriersRawACSDictionary.txt"

```{r}
acs2016 <- mutate(acs2016,
       pop16 = SE_T001_001,
       nhblk16 = SE_T014_004,
       blk16 = SE_T014_004 + SE_T014_012,
       pov16 = SE_T114_002 + SE_T115_002 + SE_T116_002,
       pov_det16 = SE_T114_001 +SE_T115_001 + SE_T116_001,
       hu16 = SE_T093_001,
       ohu16 = SE_T142_001,
       mhm16 = SE_T101_001
)
```

select needed variables from each dataset
```{r}
sub_acs16 <- select(acs2016, Geo_FIPS, pop16, nhblk16, blk16, pov16, pov_det16, hu16, ohu16, mhm16)
sub_acs16 <- rename(sub_acs16, trtid10 = Geo_FIPS)

sub_ltdb_full <- select(full1970, TRTID10, POP70, BLACK70, HU70, OHU70)
sub_ltdb_full <- rename(sub_ltdb_full, 
                       trtid10 = TRTID10,
                       pop70 = POP70,
                       black70 = BLACK70,
                       hu70 = HU70,
                       ohu70 = OHU70
)

sub_ltdb_sample <- select(sample1970, TRTID10, DPOV70, NPOV70)
sub_ltdb_sample <- rename(sub_ltdb_sample,
                         trtid10 = TRTID10,
                         pov_det70 = DPOV70,
                         pov70 = NPOV70
)
```

merge the three datasets into one, export to CSV

```{r}
cleanTractData <- merge(sub_acs16, sub_ltdb_full, by= "trtid10")
cleanTractData <- merge(cleanTractData, sub_ltdb_sample, by = "trtid10")

write.csv(cleanTractData, "/Users/thbraswell/Documents/Local/barriersCrime/cleanData/cleanTractData.csv")
```