---
title: "Street Barriers and Crime - Poisson Models"

This notebook fits poisson, negative binomial, and zero-inflated binomial models for violent- crime and street barrier locations.
---

Clear existing environment
```{r}
remove(list = ls())
```


Import packages
```{r}
# spatial packages
library(sf)
library(sp)
library(spdep)

# modeling packages
library(MASS)
library(pscl)
library(spatialreg)

# other packages
library(here)
library(stargazer)
```

Import grid data
```{r}
one_km <- st_read("../data/clean/one_km_grid.shp")
half_km <- st_read("../data/clean/half_km_grid.shp")
quarter_km <- st_read("../data/clean/quarter_km_grid.shp")
```

Calcuate mean and variance of violent crimes.
```{r}
cat("The mean number of violent crimes per one square kilometer grid square is", mean(one_km$crime), "and the variance is", var(one_km$crime), "\n\n")

```
Calcuate mean and variance of barriers.
```{r}
cat("The mean number of barriers per one square kilometer grid square is", mean(one_km$barriers), "and the variance is", var(one_km$barriers), "\n\n")

```

Get a sense of the distribution/number of zero counts for both crime and barriers.
```{r}
table(one_km$crime)
table(one_km$barriers)
```

# Spatial Weights

Create a rook spatial weights matrix
```{r}
weights <- poly2nb(one_km, queen=FALSE)
weights <- nb2listw(weights, style="C", zero.policy=TRUE)
```


## Moran's I

```{r}
moran.test(one_km$crime,weights)
```

## Crime Models
Fit a standard poisson model
```{r}
model.poisson <- glm(crime ~ scale(barriers) + scale(popE) + scale(storef)+scale(p_blk16)+scale(p_pov16)+scale(p_pooh16)+scale(p_vac16)+scale(pc50_16), data=one_km, family=poisson())
summary(model.poisson)
```


```{r}
lm.morantest(model.poisson, listw = weights)
```

We have some evidence of residual spatial autocorrelation. This suggests that we may need some kind of corrective to address this issue. Next, let's check for overdispersion. 
```{r}
AER::dispersiontest(model.poisson, alternative = "two.sided")
```
We have evidence of overdispersion, so let's move on the the negative binomial model.

## Negative Binomial

```{r}
model.nb <- glm.nb(crime ~ scale(barriers) + scale(popE) + scale(storef) + scale(p_blk16) + scale(p_pov16) + scale(p_pooh16) + scale(p_vac16) + scale(pc50_16), data=one_km)

summary(model.nb)
```

```{r}
round(exp(coef(model.nb)), 3)
```

Use the moran eigenvector filtering function to remove spatial autocorrelation from the residuals.

```{r}
me.fit<-ME(crime ~ scale(barriers) + scale(popE) + scale(storef) + scale(p_blk16) + scale(p_pov16) + scale(p_pooh16) + scale(p_vac16) + scale(pc70_16), data = one_km, family = negative.binomial(1.431), listw = weights, verbose = TRUE, alpha = .05)
```

```{r}
me.fit
```

Run a new negative binomial model with the selected eigenvectors added to the right-hand side.
```{r}
model.nb2 <- glm.nb(crime ~ scale(barriers) + scale(popE) + scale(storef) + scale(p_blk16) + scale(p_pov16) + scale(p_pooh16) + scale(p_vac16) + scale(pc50_16) + fitted(me.fit), data = one_km)

summary(model.nb2)
```

Compare spatial autocorrelation of the residuals between the negative binomial models with and without spatial eigenvector filtering.
```{r}
lm.morantest(model.nb, listw = weights)
lm.morantest(model.nb2, listw = weights)
```
We observe a minor decrease in spatial autocorrelation.

```{r}
round(exp(coef(model.nb2)), 3)
```

Write function to exponentiate the coefficients for ease of interpretation.
```{r}
coef.to.exp <- function(x) (exp(x))
```

Write function to summarize moran's i values:

```{r}
moran.summary <- function(model, weights){
  
  test <- lm.morantest(model, listw = weights)
  
  stars <- symnum(test$p.value, corr = FALSE, na = FALSE, legend = FALSE,
         cutpoints = c(0, 0.001, 0.01, 0.05), 
         symbols = c("***", "**", "*"))
  
  out <- paste0(round(as.numeric(test$estimate[1]), digits = 3), "<sup>", stars, "</sup>")
  
  return(out)
  
}
```


```{r}
stargazer(model.poisson, model.nb, model.nb2, type= "html",
          apply.coef=coef.to.exp, t.auto=F, p.auto=F, report = "vct*",
          title = "Violent Crimes per Grid Square",
          add.lines = list(
            c("AIC", round(AIC(model.poisson), digits = 3), round(AIC(model.nb), digits = 3), round(AIC(model.nb2), digits = 3)),
            c("BIC", round(BIC(model.poisson), digits = 3), round(BIC(model.nb), digits = 3), round(BIC(model.nb2), digits = 3)),
            c("Moran's I", moran.summary(model.poisson, weights), moran.summary(model.nb, weights), moran.summary(model.nb2, weights))),
          covariate.labels = c("Barriers", "Population, 2016", "Store Front Businesses", "% African American, 2016", 
                               "% Poverty, 2016", "% Owner Occupied", "% Vacant", "% Population Change, 1950-2016",
                               "Eigenvector 3", "Eigenvector 11"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          column.labels = c("(4)", "(5)", "(6)"),
          model.numbers=FALSE,
          dep.var.labels = "Violent Crimes",
          omit.stat = c("rsq", "aic"),
          digits = 3,
          out = here::here("results", "tables", "models2.html"))
```

## Count models with barriers as the dependent variable

Check for spatial autocorrelation for barriers
```{r}
moran.test(one_km$barriers,weights)
```

We're going to fit barrier models using data from 1980 to ascertain the social correlates of barriers during a time that preceded the vast majority of barrier installations.

### 1980 Data

Fit a standard poisson model
```{r}
b80_model.poisson <- glm(barriers ~ scale(pop80) + scale(p_blk80) + scale(p_vac80) + scale(p_pov80) + scale(pc50_80), data=one_km, family=poisson())

summary(b80_model.poisson)
```

Check for spatial autocorrelation in the residuals
```{r}
lm.morantest(b80_model.poisson, listw = weights)
```
Moran's I is suggestive of the presence of spatial autocorrelation in the residuals.

```{r}
AER::dispersiontest(b80_model.poisson, alternative = "two.sided")
```
Overdispersion is present, suggesting the need for a negative binomial model.

Negative binomial model

```{r}
b80_model.nb <- glm.nb(barriers ~ scale(pop80) + scale(p_blk80) + scale(p_vac80) + scale(p_pov80) + scale(pc50_80), data=one_km)

summary(b80_model.nb)
```

Use the moran eigenvector filtering function to remove spatial autocorrelation from the residuals.
```{r}
b80_me.fit<-ME(barriers ~ scale(pop80) + scale(p_blk80) + scale(p_vac80) + scale(p_pov80) + scale(pc50_80), data = one_km, family = negative.binomial(.627), listw = weights, verbose = TRUE, alpha = .05)
```

```{r}
b80_me.fit
```

```{r}
b80_model.nb2 <- glm.nb(barriers ~ scale(pop80) + scale(p_blk80) + scale(p_vac80) + scale(p_pov80) + scale(pc50_80) + fitted(b80_me.fit), data = one_km)

summary(b80_model.nb2)
```

Compare the Moran's I of the residuals
```{r}
lm.morantest(b80_model.nb, listw = weights)
lm.morantest(b80_model.nb2, listw = weights)
```

Zero-inflated model
```{r}
b80_model.zoin <- zeroinfl(barriers ~ scale(pop80) + scale(p_blk80) + scale(p_vac80) + scale(p_pov80) + scale(pc50_80), 
                           dist = "negbin", data = one_km)

summary(b80_model.zoin)
```

Zero-inflation barriers model with eigenvector filtering
```{r}
b80_model2.zoin <- zeroinfl(barriers ~ scale(pop80) + scale(p_blk80) + scale(p_vac80) + scale(p_pov80) + 
    scale(pc50_80)+ fitted(b80_me.fit) | scale(pop80) + fitted(b80_me.fit), dist = "negbin", data = one_km)

summary(b80_model2.zoin)
```

```{r}
vuong(b80_model.nb2, b80_model2.zoin)
```


```{r}
AIC(b80_model.nb)
AIC(b80_model.nb2)
AIC(b80_model.zoin)
AIC(b80_model2.zoin)
```

```{r}
stargazer(b80_model.poisson, b80_model.nb, b80_model.nb2, type = "html",
          apply.coef=coef.to.exp, t.auto=F, p.auto=F, report = "vct*",
          title = "Barriers per Grid Square",
          add.lines = list(
            c("AIC", round(AIC(b80_model.poisson), digits = 3), round(AIC(b80_model.nb), digits = 3), round(AIC(b80_model.nb2), digits = 3)),
            c("BIC", round(BIC(b80_model.poisson), digits = 3), round(BIC(b80_model.nb), digits = 3), round(BIC(b80_model.nb2), digits = 3)),
            c("Moran's I", moran.summary(b80_model.poisson, weights), moran.summary(b80_model.nb, weights), moran.summary(b80_model.nb2, weights))),
          covariate.labels = c("Population, 1980", "% African American, 1980", "% Vacant, 1980", "% Poverty, 1980", "% Population Change, 1950-1980",
                               "Eigenvector 1", "Eigenvector 2"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          dep.var.labels = "Barriers",
          omit.stat = c("rsq", "aic"),
          single.row = TRUE,
          digits = 3,
          out = here::here("results", "tables", "models1.html"))
```


```{r}
stats <- dplyr::select(one_km, barriers, crime, pop80, p_blk80, p_vac80, p_pov80, pc50_80, pc50_16, 
                popE, p_blk16, p_pov16, p_pooh16, p_vac16, storef)
st_geometry(stats) <- NULL

stargazer(stats, type = "html", 
          title = "Descriptive Statistics",
          summary.stat = c("n", "mean", "sd", "min", "max"),
          covariate.labels = c("Barrier", "Violent Crimes, 2016", "Population, 1980", "% African American, 1980",
                               "% Vacant, 1980", "% in Poverty, 1980", "% Population Change, 1950-1980",
                               "% Population Change, 1950-2016", "Population, 2016", "% African American, 2016",
                               "% in Poverty, 2016", "% Owner Occupied Homes, 2016", "% Vacant, 2016",
                               "Store Front Businesses"), 
          out = here::here("results", "tables", "descriptives.html"))
```


