---
title: "censusClean"
output: html_notebook
author: Taylor Harris Braswell
---

This file cleans raw census St. Louis tract level census data from the 2016 ACS five-year estimates and the 1970, 1980 Brown LTDB data. It will combine both data sources to create a single dataset that will be merged into a 2016 tract level shapefile for St. Louis City. 

Clear workspace
```{r}
rm(list = ls())
```

Load libraries
```{r}
library(brownLTDB)
library(tidyr)
library(dplyr)
library(sf)
```

Import raw 2016 ACS data and brown data from the LTDB libary. 

```{r}
acs2016 <- read.csv("/Users/thbraswell/Documents/Local/barriersCrime/rawData/barriersRawACS.csv", header = TRUE)
full1970 <- ltdb_stl(year = 1970, type = "full")
sample1970 <- ltdb_stl(year = 1970, type = "sample")
full1980 <- ltdb_stl(year = 1980, type = "full")
```

Wrangle ACS data. Note: raw data dictionary located in "barriersRawACSDictionary.txt"

```{r}
acs2016 <- mutate(acs2016,
       pop16 = SE_T001_001,
       nhblk16 = SE_T014_004,
       blk16 = SE_T014_004 + SE_T014_012,
       pov16 = SE_T114_002 + SE_T115_002 + SE_T116_002,
       pov_det16 = SE_T114_001 +SE_T115_001 + SE_T116_001,
       hu16 = SE_T093_001,
       ohu16 = SE_T142_001,
       mhm16 = SE_T101_001,
       hu16_4vac = SE_T095_001, # Likely, the same as hu16, but using to calculuate                                   p_vac16 after interpolation.
       vac16 =SE_T095_003
       )
```

select needed variables from each dataset
```{r}
#ACS
sub_acs16 <- select(acs2016, Geo_FIPS, pop16, nhblk16, blk16, pov16, pov_det16, hu16, ohu16, mhm16, hu16_4vac, vac16)
sub_acs16 <- rename(sub_acs16, trtid10 = Geo_FIPS)

#LTDB 1970 Full data - population, black population, housing units, occupied housing units (not needed, in hindsight), vacant housing unitys
sub_ltdb_full <- select(full1970, TRTID10, POP70, BLACK70, HU70, VAC70, OHU70)
sub_ltdb_full <- rename(sub_ltdb_full, 
                       trtid10 = TRTID10,
                       pop70 = POP70,
                       black70 = BLACK70,
                       hu70 = HU70,
                       vac70 = VAC70,
                       ohu70 = OHU70 #Note this was a mistake and pulled all occupied house units instead of owner occupied units. I don't think we'll use these, but if we do, we'll need to adjust the code and interpolate hu70 and ohu70. I did not interpolate them for this iteration because we're not pulling historical change in owner occupied units, as yet.
)

#LTDB 1970 Sample Data - poverty
sub_ltdb_sample <- select(sample1970, TRTID10, DPOV70, NPOV70)
sub_ltdb_sample <- rename(sub_ltdb_sample,
                         trtid10 = TRTID10,
                         pov_det70 = DPOV70,
                         pov70 = NPOV70)
                         
#LTDB 1980 Full data - nhblk, total population, total housing units and vacant housing units

sub_ltdb_80 <- select(full1980, TRTID10, POP80, NHBLK80, HU80, VAC80)
sub_ltdb_80 <- rename(sub_ltdb_80,
                      trtid10 = TRTID10,
                      pop80 = POP80,
                      nhblk80 = NHBLK80,
                      hu80 = HU80,
                      vac80 = VAC80
)
```

merge the four datasets into one, export to CSV

```{r}
cleanTractData <- merge(sub_acs16, sub_ltdb_full, by= "trtid10")
cleanTractData <- merge(cleanTractData, sub_ltdb_sample, by = "trtid10")
cleanTractData <- merge(cleanTractData, sub_ltdb_80, by = "trtid10")


write.csv(cleanTractData, "/Users/thbraswell/Documents/Local/barriersCrime/cleanData/cleanTractData.csv")
```

TEMPORARY - WRITE JUST 1980 DATA - DELETE BEFORE POSTING PUBLIC
```{r}
write.csv(sub_ltdb_80, "/Users/thbraswell/Documents/Local/barriersCrime/cleanData/cleanTract1980.csv")
```
